---
title: "Cross Validation"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)
library(modelr)
library(mgcv)

theme_set(theme_bw() + theme(legend.position = "bottom"))
set.seed(1)
```

## Example

Simulate data
```{r}
nonlin_df = tibble(
  id = 1:100,
  x = runif(100, 0, 1),
  y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3) # nonlinear function of x
)

ggplot(nonlin_df, aes(x = x, y = y)) + geom_point() + theme_bw()
```

Partition into training and testing

```{r}
# ?sample_n

# create the training dataset with 80 samples in the dataframe 
# (the samples are not put in order)
train_df = sample_n(nonlin_df, 80)

# put the samples left in the test dataset
# give me everything from nonlin_df dataset that are not in the train_df dataset
test_df = anti_join(nonlin_df, train_df, by = "id")

ggplot(train_df, aes(x = x, y = y)) + 
  geom_point() + 
  geom_point(data = test_df, color = "red")

# we will the black points to fit the model and see how the model predict from the point of red
```

Fit a few models...

```{r}
# linear model fit
lin_mod = lm(y ~ x, data = train_df)

# gam fits smooth curves through lines
nonlin_mod = mgcv::gam(y ~ s(x), data = train_df)

# not a good idea
wiggly_mod = mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df)
```

Show what these look like:

```{r}
# not complex enough
train_df %>% 
  # add the prediction to the dataframe
  add_predictions(lin_mod) %>% 
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

```{r}
# this is pretty good
train_df %>% 
  add_predictions(nonlin_mod) %>% 
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

```{r}
# this going out of the way to fit, not a great fit. Too complex
train_df %>% 
  add_predictions(wiggly_mod) %>% 
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

Make all the plots toghter
```{r}
train_df %>% 
  gather_predictions(lin_mod, nonlin_mod, wiggly_mod) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_wrap(~model)
```

Root-mean-squared-error

We are interested in root-mean-squared-error in test dataset

```{r}
# ?rmse
rmse(lin_mod, test_df)
## [1] 0.7163422
rmse(nonlin_mod, test_df)
## [1] 0.2437012
rmse(wiggly_mod, test_df)
## [1] 0.3471883
```

### Iterate!!

Repeat the data for bunch of random partitions

```{r}
# generate a bunch of partitions: give the dataset and number of the time you want to repeat the training test split, and each time do it at random, by fault to have 80% of the dataset for traning date, 20% of the datase for test data
cv_df = 
  crossv_mc(nonlin_df, 100) 
```

```{r}
cv_df %>% pull(train) %>% .[[1]] %>% as_tibble()
cv_df %>% pull(test) %>% .[[5]] %>% as_tibble()
```

```{r}
# change the train and test set to ba a tibble
cv_df =
  cv_df %>% 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble))
```

```{r}
cv_df = 
  cv_df %>% 
  mutate(lin_mod = map(train, ~lm(y ~ x, data = .x)),
         nonlin_mod = map(train, ~mgcv::gam(y ~ s(x), data = .x)),
         wiggly_mod = map(train, ~mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = .x))) %>% 
  mutate(rmse_lin = map2_dbl(lin_mod, test, ~rmse(model = .x, data = .y)),
         # for the rmse funtion, we have two inputs: fitted model and dataset, so use "map2"
         # use "maps_dbl" to get the number 
         rmse_nonlin = map2_dbl(nonlin_mod, test, ~rmse(model = .x, data = .y)),
         rmse_wiggly = map2_dbl(wiggly_mod, test, ~rmse(model = .x, data = .y)))
```

Summarize results...

```{r}
cv_df %>% 
  select(.id, starts_with("rmse")) %>% 
  ## tidy the data before make a plot
  gather(key = model, value = rmse, rmse_lin:rmse_wiggly) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

### Child growth

```{r}
child_growth = read_csv("./data/nepalese_children.csv")
skimr::skim(child_growth)
```

```{r}
child_growth %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = .3)
```

Is this nonlinear?
```{r}
child_growth =
  child_growth %>% 
  mutate(weight_sp = (weight > 7) * (weight - 7))
```

```{r}
lin_mod = lm(armc ~ weight, data = child_growth)
# piecewise linear model 
pwl_mod = lm(armc ~ weight + weight_sp, data = child_growth)
nonlin_mod = gam(armc ~ s(weight), data = child_growth)
```

Plot all model fits ...

```{r}
child_growth %>% 
  gather_predictions(lin_mod, pwl_mod, nonlin_mod) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red") + 
  facet_grid(~model)
```

Constructing training / testing splits

```{r}
cv_df = 
  crossv_mc(child_growth, 100) %>% 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble))
```

```{r}
cv_df = 
  cv_df %>% 
  mutate(lin_mod = map(train, ~lm(armc ~ weight, data = .x)),
         pwl_mod = map(train, ~lm(armc ~ weight + weight_sp, data = .x)),
         nonlin_mod = map(train, ~gam(armc ~ s(weight), data = as_tibble(.x)))) %>% 
  mutate(rmse_lin    = map2_dbl(lin_mod, test, ~rmse(model = .x, data = .y)),
         rmse_pwl = map2_dbl(pwl_mod, test, ~rmse(model = .x, data = .y)),
         rmse_nonlin = map2_dbl(nonlin_mod, test, ~rmse(model = .x, data = .y)))

```

Look at results..

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```


```{r}
# to double check if your predict model is doing better than absolutly nothing
cv_df = 
  cv_df %>% 
  mutate(int_mod = map(train, ~lm(armc ~ 1, data = .x)),
         lin_mod = map(train, ~lm(armc ~ weight, data = .x)),
         pwl_mod = map(train, ~lm(armc ~ weight + weight_sp, data = .x)),
         nonlin_mod = map(train, ~gam(armc ~ s(weight), data = as_tibble(.x)))) %>% 
  mutate(rmse_int    = map2_dbl(int_mod, test, ~rmse(model = .x, data = .y)),
         rmse_lin    = map2_dbl(lin_mod, test, ~rmse(model = .x, data = .y)),
         rmse_pwl = map2_dbl(pwl_mod, test, ~rmse(model = .x, data = .y)),
         rmse_nonlin = map2_dbl(nonlin_mod, test, ~rmse(model = .x, data = .y)))

cv_df %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
# it's great because at very least, the model you predict is much better than doing absolutly nothing
```